# My first knitr document 
B.H.Thu  

# Introduction
## Introduction 
### Introduction
#### Introduction

Here is the simulation of the code. Here is a code chunk 
```{r simulation, echo = TRUE, results = "hide"}
set.seed(1)
x <- rnorm(100)
mean(x)
```



```{r computetime, echo = FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```
The current time is `r time`. My favorite random number is `r rand`. 



Let's first simulate some data. 
``` {r simulatedata, echo = TRUE}
x <- rnorm(100); 
y <- x + rnorm(100, sd = 0.5)
```
Here is a scatterplot of the data 
```{r scatterplot, fig.height = 4}
par(mar = c(5,4,1,1), las = 1)
plot(x, y, main = "My simulated data")
```



Regression model
```{r fitmodel}
library(datasets)
data(airquality)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
```
Here is a table of regression coefficients
```{r showtable, results = "asis"}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type = "html")
```


Setting global options
```{r setoptions, echo = FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, results = "hide")
```

First simualate data 
```{r simulatedata1, echo = TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd = 0.5)
```
Here is a scatterplot of the data, again
```{r scatterplot1, fig.height = 4}
par(mar = c(5,4,1,1), las = 1)
plot(x, y, main = "My simulated data")
```

